---
title: "Installing Packages"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: html_document
---

Let's use `install.packages` to install a new package
```{r}
#install.packages('dadjokeapi')

# the library function loads an installed package into the environment
library(dadjokeapi)
library(dplyr)
library(tidyr)
```

Install the 'vegan' package
  * Adds many new functions to R
  * We are just going to use a dataset included with it
```{r}
#install.packages('vegan')

# load the vegan package into the environment
library(vegan)
```

Bring the "BCI" dataset into the R environment
  * This is a dataset of 50 forest plot counts of different...
  * These forest plots are like the ones you used in BIOL ...
```{r}
data(BCI)
```

This is the classic R way to calculate the mean abundance ...
  * The `$` indicates you want to refer to a particular column
```{r}
mean(BCI$Cecropia.insignis)
```

The `summarise_all` function applies a function to all columns within a dataset
  * In this case, we are applying the mean function to the...
  * The output will be the mean count of each species across...
```{r}
summarise_all(BCI,mean)
```

We can also do this with an operation called a pipe.
  * This takes an object in the line preceding it and "pipes" it into whatever is happening in the next line
```{r}
BCI2 <- BCI %>%
  summarise_all(mean) %>%
  gather()  # gathers a spread-out dataset into a tight format
head(BCI2)
```

The `sd` function calculates standard deviation

The `median` function calculates the median
