---
title: "Making Graphs in R"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
editor_options: 
  chunk_output_type: console
---

First load the vegan, tidyr, and dplyr packages
```{r}
library(dplyr)
library(tidyr)
library(vegan)
data(BCI)
```

Next, we will use a new package for making graphs, ggplot2
```{r}
library(ggplot2)
```

Let's make a histogram
* We'll use the function `ggplot`
* ggplot asks us to define "aesthetics" which are the x and y variables
* we then add to the empty `ggplot` object
* `geom_histogram` plots the probability distribution
* `xlab` and `ylab` define the axis labels
* `theme_classic` removes chartjunk
```{r}
gustavia <- BCI %>%
  ggplot(aes(x=Gustavia.superba, y=..density..)) +  # density is not an actual variable in our dataset, this is so ggplot knows to plot the histogram as density (intead of frequency, for example)
  geom_histogram() +
  xlab("Trees per Hectare") +
  ylab("Probability") +
  theme_classic()
gustavia
```

Calculate the mean and median so we can add them to the plot
```{r}
mean.gustavia = mean(BCI$Gustavia.superba)
median.gustavia = median(BCI$Gustavia.superba)
y=0
```

Add some points to the plot using `geom_point`
```{r}
gustavia +
  geom_point(aes(x=mean.gustavia, y=y), size=5, color="red") +
  geom_point(aes(x=median.gustavia, y=y), size=5, color="blue")
```

Now let's make a boxplot
* `geom_boxplot` creates a boxplot
```{r}
BCI %>%
  ggplot(aes(y=Gustavia.superba)) +
  geom_boxplot() +
  xlab("Gustavia superba") +
  ylab("Trees per Hectare") + 
  theme_classic()
```
  * Note this plot has points beyond the whiskers for outliers
  * These are >1.5x the interquartile range (the width of the box)
  * This is nice, but it only has one species
  * To make a nice boxplot with multiple categories, we need our ...
  * The `BCI` dataset is currently in the "wide format"
  * Wide format means that each observation is represented in its own column (ever row is its own observation)
  * The wide format is sometimes useful, but not usually in this context
  * We prefer the "long format," where each observation gets its...
  * Let's use `gather` to make a new object in long format
```{r}
BCI.long <- BCI %>%
  gather()

BCI.long %>%
  ggplot(aes(x=key, y=value)) +
  geom_boxplot() +
  xlab("Species") +
  ylab("Trees per Hectare") +
  theme_classic()
# way too many boxes!

# filter down
BCI.long %>%
  filter(key=="Gustavia.superba") %>%
  ggplot(aes(x=key, y=value)) +
  geom_boxplot() +
  xlab("Species") +
  ylab("Trees per Hectare") +
  theme_classic()
```
  * We can use this code to make a graph of any set species
