---
title: "Week 5 Problem Set"
author: "Emily K. Power"
date: "`r Sys.Date()`"
output: 
    prettydoc::html_pretty:
        theme: hpstr
        highlight: github
editor_options: 
  chunk_output_type: console
---

Load in libraries
```{r}
library(dplyr)
library(ggplot2)
library(ggmosaic)
library(colorBlindness)
```


1. Using the dataset collected on our class, we want to test **the null hypothesis that there is no relationship between being a Feb and being an Environmental Studies STEM major.** Would these data meet the assumptions of a chi-square test as applied to STEM majors at Middlebury College? State each assumption and what evidence you have that it is met or not.
```{r}
# load in data
PS5_data <- read.csv("/Users/emilypower/Desktop/BIOL_0211A/Data/ProblemSet5Data.csv")

# perform chi square test
chitest_PS5 <- chisq.test(PS5_data$Major, PS5_data$GradDate)

# pull expected values from test
chitest_PS5$expected
```

  * X^2 Assumptions:
    * Samples have to be randomly sampled from their respective population: This was not a random sample because the data was collected from the BIOL 0211 class alone. If we are trying to determine if there is a relationship between being a feb at Middlebury College and being an environmental STEM major among STEM majors at Middlebury, we would have to collect a random sample of STEM majors at Middlebury, rather than within a specific class. This class has prerequisites and appeals to biology students, but we are interested in more than biology students. 
    * *Less than 20% of cells have an expected frequency less than 5: Two of the cells have an expected frequency less than 5 (ENVS Feb = 2.2, Not ENVS Feb = 2.8). This is 50% of the 4 cells. Therefore, these data do not meet this assumption.
    * No cells have an expected frequency of less than 1: This is true. All cells have expected frequencies of 2.2 or above.


2. Draw a conclusion and provide the full test result (test statistic, degrees of freedom, and p value) using the most powerful and appropriate test for the hypothesis test described in Question 1. Also include the code you used to calculate this result.
```{r}
# since these data don't meet the assumptions of a chi squared test, I will perform a Fishers Exact Test
test(PS5_data$GradDate,PS5_data$Major)
```
  * We failed to reject the null hypothesis because p>0.05 and thus there is no significant relationship between being a Feb and being an environmental STEM major. 
